---
title: "statfinalpart_1"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

Final regression model: 
Int3 <- lm(price ~ landsqft + grosssqft + yrbuilt_bucket + bldclasssale*factor(neigh_cluster) + year+ quarter,data = data)

- Data coverage : 14048 observations used for regression
- Explanatory power : adjusted R^2 = 0.8774 
- Predictive power : RMSE = 253174.1
- Model parsimony : 36 model df


#1.1 Bring the data into R 
```{r}
# Load necessary library
library(tidyr)

# Set the working directory (modify this path according to your file locations)
setwd("/Users/colleenjung/Desktop/UChicago/Fall ADSP 31007 Stat Analysis")

# Load the data and assign standardized column names
B16 <- read.csv("2016_brooklyn.csv", skip=4) # Adjust this if the header row is not the first 
B17 <- read.csv("2017_brooklyn.csv", skip=4)
B18 <- read.csv("2018_brooklyn.csv", skip=4)
B19 <- read.csv("2019_brooklyn.csv", skip=4)
B20 <- read.csv("2020_brooklyn.csv", skip=6)
B20 = B20[-1,]
#View(B2020)
```


#1.2 Join the data and make it usable for analysis 
```{r}
# Define the new column names
column_names <- c('borough', 'neighborhood', 'bldclasscat', 'taxclasscurr', 'block', 'lot', 'easement',
                  'bldclasscurr', 'address', 'aptnum', 'zip', 'resunits', 'comunits', 'totunits', 
                  'landsqft', 'grosssqft', 'yrbuilt', 'taxclasssale', 'bldclasssale', 'price', 'date')

# List of dataframes
datasets <- list(B16, B17, B18, B19, B20)

# Loop through the list and change column names
for (i in seq_along(datasets)) {
  colnames(datasets[[i]]) <- column_names
}

# If the datasets are not in a list and you want to assign the names directly:
colnames(B16) <- column_names
colnames(B17) <- column_names
colnames(B18) <- column_names
colnames(B19) <- column_names
colnames(B20) <- column_names
#View(datasets)
```
```{r}
df <- rbind(B16, B17, B18, B19, B20)

# Filter the data
filtered_df <- df %>%
  filter(grepl("^A|^R", bldclasscurr),  # Starts with 'A' or 'R'
         totunits == 1,                  # 'totunits' equals 1
         resunits == 1,                     # 'units' equals 1
         grosssqft > 0,                       # Gross square footage > 0
         !is.na(price),
         price != "-")

# Print the number of rows in the filtered data
nrow(df)
nrow(filtered_df)

#filtered_df <- subset(filtered_df, price != 0)
```


#1.3 Filter the data and make transformations specific to this analysis 
```{r}
data=filtered_df
# Assuming 'price' might have non-numeric characters like '$' or ','
data$price <- gsub("\\$", "", data$price) # Remove dollar signs
data$price <- gsub(",", "", data$price)  # Remove commas
data$price <- trimws(data$price)  # Remove spaces
data$bldclasssale <- trimws(data$bldclasssale)  # Remove spaces
```
```{r}
#change to numeric data
data$totunits = as.numeric(data$totunits)
data$resunits = as.numeric(data$resunits)
data$grosssqft = as.numeric(data$grosssqft)
data$price = as.numeric(data$price)
data$borough = as.numeric(data$borough)
data$zip = as.numeric(data$zip)
#data$taxclasssale = as.numeric(data$taxclasssale)
#data$lot = as.numeric(data$lot)  # this is tax lot
data$landsqft = as.numeric(data$landsqft)
#data$date <- as.Date(data$date, format="%Y-%m-%d")


# Ensure yrbuilt is numeric
data$yrbuilt <- as.numeric(data$yrbuilt)

# Define the breaks for the year intervals
breaks <- c(1850, 1900, 1950, 1980, 2000, 2020)

# Create a new factor variable with the defined intervals
data$yrbuilt_bucket <- cut(data$yrbuilt, breaks, 
                        labels = c("1850-1900", "1901-1950", "1951-1980","1981-2000", "2001-2020"),
                        include.lowest = TRUE)
```
```{r}
#make 'year','month','quarter' columns
data$date <- as.Date(data$date, format="%m/%d/%y")
# Extract year from 'date'
data$year <- as.numeric(format(data$date, "%y"))

# Extract month from 'date'
data$month <- as.numeric(format(data$date, "%m"))

# Define the breaks for the quarters
month_breaks <- c(1, 4, 7, 10, 13)  # The 13 is to include December in Q4

# Create a new factor variable with the defined intervals for quarters
data$quarter <- cut(data$month, 
                    breaks = month_breaks, 
                    labels = c("Q1", "Q2", "Q3", "Q4"),
                    right = FALSE,  # This ensures the intervals are [), which is what you want for months
                    include.lowest = TRUE)

# Now combine with the year to create a 'Year-Q#' format
data$quarter <- quarters(data$date)




# Check the first few entries of the new variable
#head(data$year_bucket)

#View(data)
```


##Step 2: EDA and feature engineering 
#2.1 Exploratory data analysis 
```{r}
library(lattice)
#install.packages("reshape2")
library(reshape2)
```
```{r}
options(scipen = 100)
# Assuming 'filtered_dataset' is your cleaned and filtered dataset

# Histogram of Price
hist(data$price, main = "Distribution of Housing Prices", xlab = "price")

# Boxplot for Price by Neighborhood
boxplot(price ~ neighborhood, data = data, main = "Housing Prices by Neighborhood", xlab = "neighborhood", ylab = "price")

# Scatter Plot for Price vs. Gross Square Footage
plot(data$grosssqft, data$price, main = "Price vs. Gross Square Footage", xlab = "grosssqft", ylab = "price")


```
#2.2 Pre-modeling and feature engineering 
```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
library(car) # for diagnostic plots
library(Metrics) # for RMSE calculation
```

# Perform hierarchical clustering on the price vs neighborhood
```{r}
# Load necessary libraries
library(dplyr)

# Assuming your data is in a dataframe called data
# and has columns data$neighborhood and data$price

# Perform hierarchical clustering on the price column
# First, we create a matrix of the price data
prices <- matrix(data$price, ncol=1)

# Standardizing the prices (optional, but can be beneficial)
prices_standardized <- scale(prices)

# Calculating the distance matrix
distance <- dist(prices_standardized)

# Performing hierarchical clustering using complete linkage
hc <- hclust(distance, method = "complete")

# Plotting the dendrogram
plot(hc, hang = -1)

# If you decide to cut the dendrogram into clusters:
# cut the dendrogram at a height that results in the desired number of clusters
clusters <- cutree(hc, h=4) # for example, cut at height 5

# Add cluster assignments back to the original dataframe
data$neigh_cluster <- clusters

# Now you can examine the clusters
for(i in unique(data$cluster)){
  cat("Cluster", i, ":\n")
  print(table(data$neighborhood[data$cluster == i]))
  cat("\n")
}

table(data$neigh_cluster)

```
```{r}
# Assuming your data is already including the neigh_cluster assignment
profile <- data %>%
  group_by(neigh_cluster) %>%
  summarise(neighborhoods = toString(unique(neighborhood)))%>%
  filter(neigh_cluster==2)

# View the profile data frame to see the neighborhoods in each cluster
profile
#View(profile)
```

#bldclasssale: describe a property’s constructive use.
#fit; price ~ landsqft+grosssqft + yrbuilt_bucket + bldclasssale  + factor(neigh_cluster)+ year + quarter, data = data
```{r}
# Fit the linear model with the new 'quarter' variable and log-transformed 'price'
lm_model_quater <- lm(price ~grosssqft + yrbuilt_bucket + bldclasssale* factor(neigh_cluster)+ year + quarter, data = data)

# Check the summary of the model to see how well the 'quarter' variable explains variance in 'price'
summary(lm_model_quater)
```
```{r}
# Calculate RMSE
rmse_value <- sqrt(mean(lm_model_quater$residuals^2))
rmse_value
```
summary for
lm(price ~ landsqft + grosssqft + yrbuilt_bucket  + bldclasssale + neighborhood + quarter)

#Assess Model Assumptions
```{r}
# Residual vs Fitted values for Homoscedasticity
# Normal Q-Q Plot for Normality of Residuals
# Scale-Location Plot for Homoscedasticity
# Residuals vs Leverage Plot to check for influential points

#par(mfrow=c(2,2))
plot(lm_model_quater)
```

#Feature Engineering 
```{r}
#Adding Interaction Terms: for example, the effect of bldclasssale might differ depending on the neigh_cluster

library(dplyr)
data <- data %>% filter(price > 0)

#interaction_term <- bldclasssale * neigh_cluster
Int3 <- lm(price ~  log(grosssqft) + yrbuilt_bucket + factor(bldclasssale)*factor(neigh_cluster) + year+ quarter,data = data)
```
#2.3 Reach a stopping point 
```{r}
# Calculate RMSE 
rmse_Int3 <- sqrt(mean(Int3$residuals^2))
rmse_Int3
```
```{r}
summary(Int3)
```

```{r}
plot(lm_model_quater)

plot(Int3)
```

# Question:how Brooklyn home purchase prices changed between 2020,Q3 and Q4

```{r}
# Step 1: Filter out the specific levels from bldclasssale
data_filtered <- subset(data, !bldclasssale %in% c("B2", "B3", "RR"))

# Step 2: If necessary, relevel the factor (this may not be needed if step 1 removes all instances)
data_filtered$bldclasssale <- factor(data_filtered$bldclasssale)

# Fit the model again with the filtered data if necessary
# Int3_filtered <- lm(price ~ landsqft + grosssqft + yrbuilt_bucket + bldclasssale*factor(neigh_cluster) + year + quarter, data = data_filtered)

# Predict prices using the original Int3 model (assuming it was fit with all the necessary levels)
# Filter the data to Q3 and Q4 of 2020 again
data_Q3_Q4_2020_filtered <- subset(data_filtered, year == 20 & (quarter == "Q3" | quarter == "Q4"))

# Step 3: Predict housing sale prices using the Int3 model
data_Q3_Q4_2020_filtered$predicted_price <- predict(Int3, newdata = data_Q3_Q4_2020_filtered)

# Step 4: Plot the predicted prices for Q3 and Q4
library(ggplot2)

ggplot(data_Q3_Q4_2020_filtered, aes(x = quarter, y = predicted_price)) +
  geom_boxplot() +
  labs(title = "Predicted Housing Sale Prices in Q3 and Q4 of 2020 (Filtered Data)",
       x = "Quarter",
       y = "Predicted Price") +
  theme_minimal()+
  scale_y_continuous(limits = c(0, 10000000))

```
#Density Plot
```{r}
ggplot(data_Q3_Q4_2020_filtered, aes(x = predicted_price, color = quarter)) +
  geom_density(alpha = 0.7) +
  labs(title = "Density of Predicted Housing Sale Prices in Q3 and Q4 of 2020",
       x = "Predicted Price",
       y = "Density") +
  theme_minimal()+ 
  scale_x_continuous(limits = c(0, 7500000))

```
#Histogram
```{r}
ggplot(data_Q3_Q4_2020_filtered, aes(x = predicted_price, fill = quarter)) +
  geom_histogram(position = "dodge", bins = 30, alpha = 0.7) +
  labs(title = "Histogram of Predicted Housing Sale Prices in Q3 and Q4 of 2020",
       x = "Predicted Price",
       y = "Count") +
  theme_minimal()+
  scale_x_continuous(limits = c(0, 7500000))

```
#Boxplot
```{r}
# Filter the data to include only prices within the specified range
data_Q3_Q4_2020_filtered <- subset(data_Q3_Q4_2020_filtered, predicted_price >= 0 & predicted_price <= 10000000)

# Create the box plot
ggplot(data_Q3_Q4_2020_filtered, aes(x = quarter, y = predicted_price)) +
  geom_boxplot() +
  labs(title = "Box Plot of Predicted Housing Sale Prices in Q3 and Q4 of 2020",
       x = "Quarter",
       y = "Predicted Price") +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 10000000))

```
#Calculate Original Price Change:
```{r}
original_Q3_mean <- mean(subset(data_filtered, quarter == "Q3" & year == 20)$price, na.rm = TRUE)
original_Q4_mean <- mean(subset(data_filtered, quarter == "Q4" & year == 20)$price, na.rm = TRUE)
original_change <- (original_Q3_mean-original_Q4_mean)
original_change
```
```{r}
original_Q4_mean/original_Q3_mean
```

#Calculate Predicted Price Change
```{r}
predicted_Q3_mean <- mean(subset(data_Q3_Q4_2020_filtered, quarter == "Q3")$predicted_price)
predicted_Q4_mean <- mean(subset(data_Q3_Q4_2020_filtered, quarter == "Q4")$predicted_price)
predicted_change <- (predicted_Q3_mean-predicted_Q4_mean)
predicted_change
```
```{r}
multiplicative_change <- predicted_Q3_mean / predicted_Q4_mean
multiplicative_change
```

```{r}
comparison_data <- data.frame(
  Category = c("Original", "Predicted"),
  Q3_Mean = c(original_Q3_mean, predicted_Q3_mean),
  Q4_Mean = c(original_Q4_mean, predicted_Q4_mean),
  Change = c(original_change, predicted_change)
)

ggplot(comparison_data, aes(x = Category, y = Change, fill = Category)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Comparison of Original and Predicted Price Changes",
       x = "Category",
       y = "Percentage Change") +
  theme_minimal()

```
```{r}
# Create a dataframe for the original and predicted means for Q3 and Q4
comparison_data <- data.frame(
  Quarter = rep(c("Q3", "Q4"), 2),
  Price_Type = c(rep("Original", 2), rep("Predicted", 2)),
  Mean_Price = c(original_Q3_mean, original_Q4_mean, predicted_Q3_mean, predicted_Q4_mean)
)

comparison_data
```

